---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```
## Check package installation
```{r warning=FALSE}
libs <- c("plotly", "data.table", "knitr", "dplyr", "stringr")
for (i in libs){
  if( !is.element(i, .packages(all.available = TRUE)) ) {
    install.packages(i)
  }
  library(i,character.only = TRUE)
}
```

## Loading and preprocessing the data
```{r}
data <- read.table(unz("activity.zip", "activity.csv"), header=T, quote="\"", sep=",")
data <- data.table(data)
sm <- summary(data)
sm
```

## What is mean total number of steps taken per day?

```{r, warning=FALSE}
# Process data
StepsPerDay <- data %>% 
  group_by(date) %>% 
  summarise(Total_Steps = sum(steps, na.rm = TRUE))

# Plot
compute_bins <- function(x, n) {
    list(
      start = min(x),
      end = max(x),
      size = (max(x) - min(x)) / n
    )
}

p1 <- plot_ly(data = StepsPerDay,
                 x = ~Total_Steps,
              type = "histogram",
          autobinx = FALSE,
             xbins = compute_bins(StepsPerDay$Total_Steps, 20)
             ) 
p1
MeanSteps <- round(mean(StepsPerDay$Total_Steps), 2)
MedianSteps <- median(StepsPerDay$Total_Steps)
```
Mean total number of steps taken per day: `r MeanSteps`  
Median total number of steps taken per day: `r MedianSteps`

## What is the average daily activity pattern?
```{r warning=FALSE}
# Function of processing
processing <- function(data, ...) {
  dots <- as.character(list(...))
  # Process data
  MeanStepsPerInterval <- data 
  MeanStepsPerInterval <- data.table(MeanStepsPerInterval)
  MeanStepsPerInterval[, Average_Steps := mean(steps, na.rm = TRUE), by = dots]
    
  # Convert interval to time
  MeanStepsPerInterval[, Time := str_pad(interval, width = 4, side = "left", pad = "0")]
  MeanStepsPerInterval$Time <- paste0(substr(MeanStepsPerInterval$Time, start = 1, stop = 2),
                                      ":",
                                      substr(MeanStepsPerInterval$Time, start = 3, stop = 4)
                                      )
  MeanStepsPerInterval[, c("steps", "date", "interval") := NULL]
  
  unique(MeanStepsPerInterval)
}

MeanStepsPerInterval <- processing(data, "interval")

# Annotation
m <- MeanStepsPerInterval[which.max(MeanStepsPerInterval$Average_Steps), ]
annotation <- list(x = m$Time,
                   y = m$Average_Steps,
                text = paste("Maximum number of steps in 5-minute interval</br></br>Time:", 
                             m$Time,
                             "</br>Average Steps:",
                             round(m$Average_Steps))
                   )
# Plot
plot_ly(data = MeanStepsPerInterval,
        x = ~Time,
        y = ~Average_Steps,
        type = 'scatter',
        mode = 'line',
   hoverinfo = 'text',
        text = ~paste('Time: ', Time, '</br></br>Average steps: ', round(MeanStepsPerInterval$Average_Steps))
   ) %>% layout(annotations = annotation)
```


## Imputing missing values
In `summary(data)`, we can know that there are `r sm[nrow(sm),1]` missing value in column `steps`, so we can set the missing value as the mean value of total steps per day.  
```{r}
StepsPerDayImputed <- StepsPerDay
StepsPerDayImputed[StepsPerDay$Total_Steps == 0, ]$Total_Steps <- mean(StepsPerDay$Total_Steps)

p2 <- plot_ly(data = StepsPerDayImputed,
                 x = ~Total_Steps,
              type = "histogram",
          autobinx = FALSE,
             xbins = compute_bins(StepsPerDay$Total_Steps, 20)
             ) 

p2
MeanStepsImputed <- round(mean(StepsPerDayImputed$Total_Steps), 2)
MedianStepsImputed <- median(StepsPerDayImputed$Total_Steps)
```
Mean total number of steps taken per day: `r MeanStepsImputed`  
Median total number of steps taken per day: `r MedianStepsImputed`  


## Are there differences in activity patterns between weekdays and weekends?
```{r}
# Convert the date to weekday or weekend
data$dateType <- ifelse(as.POSIXlt(data$date)$wday %in% c(0,6), 'Weekend', 'Weekday')

MeanStepsPerIntervalImputed <- processing(data, "interval", "dateType")
# Plot
plot_ly(data = MeanStepsPerIntervalImputed,
                 x = ~Time,
                 y = ~Average_Steps,
             color = ~dateType,
              type = "scatter",
              mode = "line"
             )
```

